---
title: "Class11: Structural Bioinformatics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB database

The [PDB](http://www.rcsb.org/) is the main repository for biomolecular structure data.

Here we examine the contents of the PDB:

```{r}
db<-read.csv("Data Export Summary.csv", row.names = 1)
db
```

How many are X-Ray
```{r}
(db$Total[1]/sum(db$Total))*100
```

How many are X-Ray, etc...
```{r}
(db$Total/sum(db$Total))*100
```

What proportion of structures are protein?
```{r}
(sum(db$Proteins)/sum(db$Total))*100
```

We could also try the datapasta package and copy from website and "Addins" > "Paste as data.frame"
```{r}
library(datapasta)

temp<- data.frame (stringsAsFactors=FALSE,
   Experimental.Method = c("X-Ray", "NMR", "Electron Microscopy", "Other",
                           "Multi Method", "Total"),
              Proteins = c(126880, 11062, 2277, 256, 129, 140604),
         Nucleic.Acids = c(2012, 1279, 31, 4, 5, 3331),
    ProteinComplex = c(6547, 259, 800, 6, 2, 7614),
                 Other = c(8, 8, 0, 13, 1, 30),
                 Total = c(135447, 12608, 3108, 279, 137, 151579))

temp
```

>Q2.Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

There are 1157 as of 2019-05-07 See:
http://www.rcsb.org/pdb/results/results.do?tabtoshow=Current&qrid=7F9EE90A



## Introduction to bio3D

```{r}
library(bio3d)
```

```{r}
pdb<-read.pdb("1hsg")
pdb
```

```{r}
attributes(pdb)
```

```{r}
head(pdb$atom)
```

# Print a subset of $atom data for the first two atoms
```{r}
pdb$atom[1:2, c("eleno","elety", "x","y","z")]
```

# Note that individual $atom records can also be accessed like this
```{r}
pdb$atom$elety[1:2]
```

# Which allows us to do the following
```{r}
plot.bio3d(pdb$atom$b[pdb$calpha],sse=pdb,typ="l", ylab="B-factor")
```

# Print a summary of the coordinate data in $xyz
```{r}
pdb$xyz
```

# Examine the row and column dimensions
```{r}
dim(pdb$xyz)
```

# Print coordinates for the first two atom
```{r}
pdb$xyz[ 1, atom2xyz(1:2) ]
```

# Atom selection is done via the function **atom.select()** 
```{r}
inds<-atom.select(pdb, "protein")
```


```{r}
prot.pdb<-atom.select(pdb, "protein",value=TRUE)
write.pdb(prot.pdb, file="1hsg_protein.pdb")
```

```{r}
lig.pdb<-atom.select(pdb,"ligand",value=TRUE)
write.pdb(lig.pdb, file="1hsg_ligand.pdb")
```

## Section 5

```{r}
aa<- get.seq("1ake_A")
```

```{r}
b <- blast.pdb(aa)
```

