---
title: 'Class9: Unsupervised Learning Mini-Project'
author: "Victoria Shi"
date: "4/30/2019"
output: github_document
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Unsupervised Learning Analysis of Human Breast Cancer Cells

## Exploratory data analysis
```{r}
fna.data<- read.csv("https://bioboot.github.io/bimm143_S19/class-material/WisconsinCancer.csv")
```

```{r}
wisc.df<-fna.data
wisc.df
```
Q: How many patients are in this data?
```{r}
nrow(wisc.df)
```


Q: How many of the observations have a malignant diagnosis?
```{r}
table(wisc.df$diagnosis)
```


```{r}
wisc.data<- as.matrix(wisc.df[3:32])
```

```{r}
row.names(wisc.data)<-wisc.df$id
head(wisc.data)
```

>Q: How many of the observations have a malignant diagnosis?
```{r}
diagnosis<-as.numeric(wisc.df$diagnosis=="M")
diagnosis
```

>Q:How many variables/features in the data are suffixed with _mean?
```{r}
colnames(wisc.data)
```

```{r}
length(grep("_mean",colnames(wisc.data)))
colnames(wisc.data)
```

## Principal Component Analysis
```{r}
round(colMeans(wisc.data,1))
```

```{r}
round(apply(wisc.data,2,sd),1)
```

```{r}
wisc.pr<-prcomp(wisc.data,scale=TRUE)
summary(wisc.pr)
```

```{r}
biplot(wisc.pr)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=diagnosis+1)
```

```{r}
pr.var<-(wisc.pr$sdev^2)
head(pr.var)
```

```{r}
pve<-pr.var/(sum(pr.var))*100
```

```{r}
plot(pve,xlab="Principal Component",ylab="Proportion of Variance Explained", ylim=c(0,100), type="o")
```

```{r}
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```

```{r}
plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 100), type = "o")
```

```{r}
par(mfcol=c(1,2))

plot(pve,xlab="Principal Component",ylab="Proportion of Variance Explained", ylim=c(0,100), type="o")

plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 100), type = "o")
```

## Hierarchical clustering

```{r}
data.scaled<- scale(wisc.data)
```

```{r}
data.dist<-dist(data.scaled)
```

```{r}
wisc.hclust<-hclust(data.dist,method="complete")
```

```{r}
plot(wisc.hclust)
abline(h=19,col="red",lty=2)
```

```{r}
wisc.hclust.clusters<-cutree(wisc.hclust,k=4)
wisc.hclust.clusters
```

```{r}
table(wisc.hclust.clusters,diagnosis)
```

## Combining methods

```{r}
data.dist2<-dist(wisc.pr$x[,1:7])
```

```{r}
wisc.pr.hclust<-hclust(data.dist2,method="ward.D2")
```

```{r}
plot(wisc.pr.hclust)
abline(h=19,col="red",lty=2)
```

```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```

```{r}
table(grps, diagnosis)
```

```{r}
plot(wisc.pr$x[,1:2], col=grps)
```

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis+1)
```

```{r}
library(rgl)
plot3d(wisc.pr$x[,1:3], xlab="PC 1", ylab="PC 2", zlab="PC 3", cex=1.5, size=1, type="s", col=diagnosis+1)
```

```{r}
library(rgl)
plot3d(wisc.pr$x[,1:3], xlab="PC 1", ylab="PC 2", zlab="PC 3", cex=1.5, size=1, type="s", col=diagnosis+1)
rglwidget(width = 400, height = 400)

```


```{r}
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=2)
```

```{r}
table(wisc.pr.hclust.clusters, diagnosis)
```

```{r}
table(wisc.hclust.clusters, diagnosis)
```


## Prediction

```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1:2], col=grps)
points(npc[,1], npc[,2], col="blue", pch=16)
```

