---
title: 'Class8: Machine learning'
author: "Victoria Shi"
date: "4/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering

Let's start with an example of running the **kmeans()** function

```{r}
 # Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```



Use the kmeans() function setting k to 2 and nstart=20
Inspect/print the results
>Q. How many points are in each cluster?
30
>Q. What ‘component’ of your result object details
      - cluster size?
      - cluster assignment/membership?
      - cluster center?
>Plot x colored by the kmeans cluster assignment and
      add cluster centers as blue points
      
```{r}
cluster<-kmeans(x, centers = 2, nstart = 20)
cluster
```

```{r}
cluster$size
```

```{r}
cluster$centers
```

```{r}
cluster$cluster
```

```{r}
plot(x,col=cluster$cluster)
points(cluster$centers, pch=20, col="blue", cex=3)
```


## Hierarchical Clustering example

We must give the **hclust()** function a distance matrix not the raw data as input

```{r}
# Distance matrix calculation
d<-dist(x)

# Clustering
hc<-hclust(d)
plot(hc)
```

```{r}
plot(hc)
abline(h=6,col="red")
```

```{r}
cutree(hc,h=6)
```

```{r}
cutree(hc, k=2)
```

```{r}
cutree(hc, k=3)
```



```{r}
 # Step 1. Generate some example data for clustering
x <- rbind(
  matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2),   # c1
  matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
  matrix(c(rnorm(50, mean = 1, sd = 0.3),           # c3
           rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
#         (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

>Q. Use the dist(), hclust(), plot() and cutree() functions to return 2 and 3 clusters
>Q. How does this compare to your known 'col' groups?

```{r}
hc<-hclust(dist(x))
plot(hc)
abline(h=2,col="red")
abline(h=2.8, col="blue")
```

```{r}
gp2<-cutree(hc,k=2)
gp3<-cutree(hc,k=3)

```

```{r}
gp2
```

```{r}
plot(x,col=gp2)
```



```{r}
gp2
```

```{r}
gp3
```

```{r}
table(gp2)
```

```{r}
table(gp3)
```

```{r}
table(gp2,gp3)
```


# PCA: Principal Component Analysis

We will use the **prcomp()** function for PCR
```{r}
## You can also download this file from the class website!
     mydata <- read.csv("https://tinyurl.com/expression-CSV",
                        row.names=1)
```

100 genes in this dataset
```{r}
nrow(mydata)
```

```{r}
ncol(mydata)
```

```{r}
colnames(mydata)
```

Running out PCA analysis on the transpose of our data.
```{r}
pca<-prcomp(t(mydata), scale=TRUE)
```

PCA plot
```{r}
## A basic PC1 vs PC2 2-D plot
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2")
```

Variance captured per PC
```{r}
pca.var <- pca$sdev^2
```

Calculate the percent variance captures in each PC
```{r}
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
```

```{r}
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
barplot(pca.var.per, main="Scree Plot",
        xlab="Principal Component", ylab="Percent Variation")
```


```{r}
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
     xlab=paste0("PC1 (", pca.var.per[1], "%)"),
     ylab=paste0("PC2 (", pca.var.per[2], "%)"))
```

#PCA of UK food data

```{r}
x <- read.csv("UK_foods.csv",row.names=1)
```

```{r}
nrow(x)
ncol(x)
```

```{r}
head(x,6)
```

```{r}
dim(x)
```

```{r}
barplot(as.matrix(x), beside=TRUE, col=rainbow(nrow(x)))
```

```{r}
barplot(as.matrix(x), beside=FALSE, col=rainbow(nrow(x)))
```

```{r}
pairs(x, col=rainbow(10), pch=16)
```

```{r}
pca<-prcomp(t(x))
summary(pca)
```

```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x),col=c("red","yellow","blue","green"))
```

```{r}
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v
```

```{r}
z <- summary(pca)
z$importance
```

```{r}
barplot(v, xlab="Principal Component", ylab="Percent Variation")
```

```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```

```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,2], las=2 )
```

```{r}
biplot(pca)
```



