---
title: "Class7: R functions and packages"
author: "Victoria Shi"
date: "4/23/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions revisited

We will source a file from online with our functions from last day

```{r}
source("http://tinyurl.com/rescale-R")
```

Try out the last day's rescale() function

```{r}
rescale(1:10)
```


Try the rescale2() function that catches string inputs
```{r}
 rescale2 <- function(x, na.rm=TRUE, plot=FALSE, ...) {
   if( !is.numeric(x) ) {
      stop("Input x should be numeric", call.=FALSE)
   }
   rng <-range(x, na.rm=na.rm)
   answer <- (x - rng[1]) / (rng[2] - rng[1])
   if(plot) {
      plot(answer, ...)
}
   return(answer)
}
```


# Find missing NA values in two vectors

Start with a simple example of the larger problem I am trying to solve
```{r}
x <- c( 1, 2, NA, 3, NA) 
y<-c(NA,3,NA,3, 4)
```

```{r}
is.na(x)
```

```{r}
is.na(y)
```

Try putting these together with an AND

```{r}
is.na(x)&is.na(y)
```

Take the sum() to find out how may TRUE values we have and thus how many NAs we had in both x and y

```{r}
sum(is.na(x)&is.na(y))
```

Now I can make this into our first function...

```{r}
both_na<-function(x,y){
  sum(is.na(x)&is.na(y))
}
```

```{r}
both_na(x,c(NA,3,NA,2,NA))
```

Test, test, test

```{r}
 x <-  c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
```

```{r}
3==3
```

```{r}
3<2
```

```{r}
3!=2
```

```{r}
length(x)
```

```{r}
length(y2)!=length(x)
```

Now let's try the both_na2() function on our different length input vectors

```{r}
which(c(F,F,T,F,T))


which(is.na(c(1,2,NA,4)))
```



```{r}
x <- c( 1, 2, NA, 3, NA) 
y<-c(NA,3,NA,3, 4)
both_na3(x,y)
```


## Intersect function

```{r}
df1
```

```{r}
df2
```

Make things simple

```{r}
x<-df1$IDs
y<-df2$IDs

x
y
```

```{r}
intersect(x,y)
```

A working snippet

```{r}
gene_intersect<-function(x,y){
cbind(x[x %in% y],
      y[y %in% x])
}
```

```{r}
gene_intersect(df1$IDs,df2$IDs)
```


```{r}
gene_intersect2(df1,df2)
```


```{r}
gene_intersect3(df1,df2)
```


```{r}
merge(df1,df2,by="IDs")
```


## Grade function

Find average grade dropping the worst homework score

```{r}
x<-c(100,100,100,100,100,100,100,90)
```

```{r}
grade <- function(x) {
  (sum(x)-min(x))/(length(x)-1)
}
```

```{r}
y<-c(100,90,90,90,90,90,97,80)
grade(y)
```


